Apparent preferences for cognitive effort fade when multiple forms of effort

### 2020-11-19 Cost_manuscript.Rmd
new discussion paragraph

  The base formulation lacked the flexibility needed to capture the dynamics observed in the within-subjects experiment. We therefore tested a second model which allowed $\gamma$ to evolve over time, adapting a learning rule from Constantino & Daw (2015)

@@

$$\begin{aligned}
mS_i = (s_i w_i) + (S_{demand} (1 - w_i))
\end{aligned}$$

@@

  Understanding the sporadic attractiveness of effortful options remains elusive. In this study, we tested the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when tasks are interleaved.

@@

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. The gradual choice convergence in the second experiment also discard the effects of boredom, which would have accentuated the appeal of active tasks over time, or the possibility that all our participants were high on need for cognition [@Cacioppo1996; @Sayali2019]. Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting [@Wikenheiser2013; @Sweiss2018], as participants expressed their choices as early as possible, resulting in occassional underharvesting. Finally, the observed differences in acceptances between effort types cannot be explained by reward alone, as demands were rewarded at the same rate.

@@

  Our foraging paradigms allowed us to control for common shortfalls in the existing literature. For example, even though engaging with a demanding offer encompasses evaluating, pursuing, and receiving an outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting subjects with choices whose results do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [@Kable2007; @Chong2016; @Massar2015; @Apps2015a]. However, discrepancies between expected and experienced value are well known [@Kahneman2006], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards, or hoping to find better offers in the future [@Fawcett2012; @Frederick2002; @Sozou1998]. Here we gave participants full knowledge of the available options, and provided an opportunity to actively experience the demands they accepted. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. This is often circumvented by comparing discounting functions and points at which individuals become indifferent to the increment in rewards [@Massar2015; @Prevost2010]. Instead, we eliminated this possibility by fixating the handling time for all demands, which allowed us to probe people's preferences directly. In addition, the between-subjects experiment provided a baseline of foraging behavior, particularly in the traditional experimental case of foraging by waiting for the next reward []. This baseline helped us interpret the findings of the second experiment. Finally, the existence of optimal strategies let us know that people acted as if effort was attractive. This was evidenced by both the lower earnings by the cognitive effort group, and by the base OC model. Of note is that the optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.

@@

- A potential concern is that our cognitive and physical demands were not phenomenologically effortful.
- Note that the calibration of efforts were different. We adjusted to people's physical capabilities, but trained them in the cognitive tasks until they reached criterion. 
- We still think that cognitive and physical manipulations were effortful. Our physical effort task has been repeatedly used to probe effort []. Similar task switching has been shown to feel cognitively effortful by recruiting cognitive control (Kool, some amitai). Furthermore, the effortful nature of our manipulations is evidenced by the steeper decrease in acceptances over time at longer handling times (Treadway). There are many reasons why such cognitive manipulations can be effortful, which are often centered in the amount of control. Or the OC. 
- Efficacy, along with reward amount, can influence how much control is exherted. AFAIK, Eisenberger1982 provides evidence that task diversity motivates further effort engagement.
- But here we do not focus on choosing the *amount* of cognitive control to be exherted during performance, which can be due to efficacy, agency, or reward amount. Instead, we focus on what makes people *pursue* the task in spite of how effortful it is.

>>>>

  The base formulation lacked the flexibility needed to capture the dynamics observed in the within-subjects experiment. We therefore tested a second model which allowed $\gamma$ to evolve over time, adapting a learning rule from @Constantino2015

@@

$$\begin{aligned}
mS_i = s_i w_i + S_{demand} (1 - w_i)
\end{aligned}$$

@@

  Understanding the sporadic attractiveness of effortful demands remains elusive. In this study, we tested the hypothesis that the perceived cost or value of a demand could depend on the ability to achieve the same outcome through different means. Using two foraging experiments that exposed individuals to a single or intermixed types of demands, we found that 1) preferences for cognitive effort were stable when individuals faced a single form of demand, but faded over time when other costs were also experienced; and 2) the phenomenon can be explained by a model in which preferences depend on how participants construe the amount of time a task takes, which could hypothetically be recalibrated when tasks are interleaved.

@@

  These choice dynamics ruled out a number of common explanatory candidates. For instance, differences in choice stability between experiments suggest that a demand's cost is not intrinsic. The gradual choice convergence in the second experiment also discard the effects of boredom, which would have accentuated the appeal of active tasks over time, or the possibility that our participants were demand seekers [@Cacioppo1996; @Sayali2019]. Even though fatigue could explain the decrease in effort acceptances across experiments, we would have expected tired participants to turn to wait trials over time in order to preserve earning rates. We also ruled out the sunk cost effects used to explain overharvesting [@Wikenheiser2013; @Sweis2018], as participants expressed their choices as early as possible, resulting in occassional underharvesting. Finally, the observed differences in acceptances between effort types cannot be explained by reward alone, as demands were rewarded at the same rate.

@@

  Our foraging paradigms allowed us to control for common shortfalls in the existing literature. For example, even though engaging with a demanding offer encompasses evaluating, pursuing, and receiving its outcome, investigations of the pursuit and receipt stages have been limited. Many studies have sought to isolate the evaluation stage by presenting individuals with offers that do not come into effect until the end of the experimental session (with delays spanning up to months), deliberately setting aside the question of how the ongoing experience of cost and outcome affects subsequent choices [@Kable2007; @Chong2016; @Massar2015; @Apps2015a]. However, discrepancies between expected and experienced value are well known [@Kahneman2006], and choosing immediate outcomes could be due to factors other than time preference, such as the risk of being unable to redeem the future rewards, or hoping to find better offers in the near future [@Fawcett2012; @Frederick2002; @Sozou1998]. We avoided these pitfalls by providing participants with an opportunity to actively experience the demands they accepted, and gave them full knowledge of the available options. Another challenge in comparing effort and delay is that increments in the required effort are often paired with a longer execution time. This is often circumvented by comparing discounting functions, and points at which individuals become indifferent to the increment in rewards [@Massar2015; @Prevost2010]. We eliminated this possibility by fixating the handling time for all demands, which allowed us to probe people's preferences directly. In addition, the between-subjects experiment provided a baseline of foraging behavior, particularly in the traditional experimental case of foraging by waiting for the next reward []. This baseline helped us interpret the findings of the second experiment. Finally, the existence of optimal strategies helped us identify when demands could be attractive, a central motivation for this study. The apparent attractiveness of cognitive effort was reflected both by a higher proportion of acceptances at the expense of earnings, and by comparisons of fitted opportunity costs against the optimal earning rate. Of note is that the optimality of the strategies produced by each cost depended on the richness of the block. This can help explain why the wait group earned the most despite having similar acceptance rates to the physical group, as the former was closer to optimality in the richest environment.

  It is possible that the task-switching configuration of the cognitive tasks was not effortful, maybe due to our calibration. Unlike adjusting to people's physical capabilities, we trained participants in the cognitive tasks until they reached criterion. Attributes such as the efficacy or capacity to complete a task, or the diversity of its requirements, have been reported to promote engagement [ivan; eisenberger]. Even so, we believe our manipulations were successful in inducing feelings of effort. Hand grips are regularly used to probe physical effort [@Chong2017; @Hogan2018; @Prevost2010], and task switching has been shown to feel cognitively effortful by recruiting cognitive control [@Kool2010; @Shenhav2017]. Furthermore, we observed a steeper decrease in acceptances over time at longer handling times, a hallmark of effortful decisions [@Treadway2009]. Regardless, our focus here was not on the amount of effort exherted, but on what makes people pursue the task in spite of its toll.

@@
